---
title: "Tidy Tuesday: UFO Encounters"
author: "Charles Davis"
date: "6/25/2019"
output:
  html_document:
    toc: yes
  html_notebook:
    code_folding: hide
    df_print: paged
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(cowplot)

setwd("~/Desktop/tidytuesday/data/2019/2019-06-25")
ufo = read.csv("ufo_sightings.csv")

```

OK. What can we ask with this dataset?

- Do the number of UFO sighting vary by country? Does the length of UFO encounter vary by country?

```{r country}

ufo %>% 
  group_by(country) %>%
  summarise(count=n())

ufo %>%
  group_by(country) %>%
  summarise(avg_length=mean(encounter_length, na.rm=TRUE))

```

So it's pretty clear that the US is the main UFO hotspot. Let's focus our attention there. Are there particularly good UFO-viewing spots in the US? 

```{r state n}

# first we need to load in some map data

state_map <- map_data("state")
stateInfo=cbind.data.frame(abb=tolower(state.abb), name=tolower(state.name))
state <- inner_join(state_map, stateInfo, by=c("region"="name"))

# let's check the counts of UFO sightings by state

state_n <- ufo %>%
  dplyr::filter(country == "us") %>%
  group_by(state) %>%
  summarise(count=n())

state_n <- inner_join(state, state_n, by=c("abb"="state"))
state_n$abb <- as.factor(state_n$abb)

ggplot(data=state, mapping=aes(x=long, y=lat, group=group)) + 
  geom_polygon(data=state_n, aes(fill=count), color="white") +
  theme_void() +
  labs(fill="number of encounters") +
  scale_fill_gradientn(colors=RColorBrewer::brewer.pal(name="Greens", n=48))

```

OK. That's kind of boring. Where more people live, more people see aliens. But maybe there are places where UFO encounters are particularly long. 

```{r state len}

ufo %>% 
  dplyr::filter(country == "us") %>% 
  ggplot(aes(x=as.factor(encounter_length))) + 
  geom_histogram(stat="count") +
  theme(axis.title.x = element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())

# since the data are clearly skewed, we'll do the rest of the analysis on log-transformed encounter length

state_len <- ufo %>%
  dplyr::filter(country == "us") %>%
  group_by(state) %>%
  summarise(avg_length=mean(log(encounter_length), na.rm=TRUE))

state_len <- inner_join(state, state_len, by=c("abb"="state"))
state_len$abb <- as.factor(state_len$abb)

length_plot <- ggplot(data=state, mapping=aes(x=long, y=lat, group=group)) + 
  geom_polygon(data=state_len, aes(fill=avg_length), color="white") +
  theme_void() +
  labs(fill="encounter length in log(min)") +
  scale_fill_gradientn(colors=RColorBrewer::brewer.pal(name="Greens", n=48)) +
  ggtitle("Length of UFO encounters across the United States") +
  theme(plot.title = element_text(size=32, face="bold"), legend.position = "bottom") 
length_plot

alien_plot <- ggdraw() + 
  draw_plot(length_plot) +
  draw_image("~/Desktop/tidytuesday/data/2019/2019-06-25/alien.jpg", scale=0.3, width=1.8, height=0.5)
alien_plot
```

Aha! As we might expect, New Mexico and Arizona are right up there. The clear skies of Maine also seem conducive to UFO sightings. 